---
title: "PADP 7120 Data Applications in PA"
subtitle: "RLab 1: Intro to R"
author: "Alex Combs"
institute: "UGA | SPIA | PADP"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    css: 'mypres.css'
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      countIncrementalSlides: false
---
# Objectives

By the end of this lab, you will learned how to...

- open and navigate R
- start a R project to organize your work
- load data already formatted for R
- load R packages
- use R markdown and run code
- export your R markdown document to HTML

---
# RLab Conventions

- Bullets denote background, instructions, etc.

- Block quote in bold like below denotes a prompt

> **Do this thing in this block**

- Example code is in blocks like below or this `in-line code`

```{r, eval=FALSE}
summary(cars)
```

---
# RLab Process

- Not recommended that you work ahead even if you are waiting for others to catch up

- Highly likely that you will encounter errors; this is normal
  - If so, **STOP ME** with your voice or a Zoom icon like raised hand or stop sign
  - Be prepared to share your screen
  - Troubleshooting can help you and others learn
  - Observers are encouraged to offer suggestions

---
# Starting R

- Once R and RStudio are installed on your computer

> **Launch RStudio**

- Should see 3 Panes
  - Console (left)
  - Environment (top-right)
  - Files and other tabs (bottom-right)

---
# File Directory

> **Type `getwd()` into the console (bottom-left pane) and hit Enter**

- Should now see a printout of the current file path; for example

- "/Users/alexcombs/Documents/"

- When you tell R to look for or save a file, this is the location R uses

- We usually want to separate folder for each project/assignment to hold all of our data and documents

---
# R Projects

- R Projects are the best way to organize your work

- Creates a folder for R to work from during the current session

--

- I will now demonstrate how to start an R Project

--

> ** Now create a project called "rlab1" in a location of your choice**

--

- All of our R Labs will begin with this process

- R Demo 1 on eLC demonstrates this process in case you forget how

---
# Loading data already formatted for R

- Certain software uses certain file extensions like `.xlsx` for Excel

- A data file that is already formatted for R has the extension `.RData`

--

> **Download the file titled "states1.RData" on eLC. Leave it wherever your computer saved it (probably your Downloads folder)**

--

> **Type `load('states1.RData')` in the console and hit `Enter`**

- Remember you must use single or double quotes anytime you tell R to do something with a file

---
# Our first error

- Should now see an error message like the one below

```{r, eval=FALSE}
Error in load(states1.RData) : object 'states1.RData' not found
Calls: <Anonymous> ... withCallingHandlers -> withVisible -> eval -> eval -> load
Execution halted
```

- Be sure to read error messages; R tries to tell you what went wrong

--

- Remember, R looks for files in your project folder

- R is saying it can't find the file we just told it to load

> **Using your computer's file manager, move states1.RData to your project folder**

> **Now try to execute the load**

---
# Data in your R session

- The environment pane (top-right) will list all data and any other objects currently loaded

- You should see `states1` in your environment pane

- You can click the dropdown arrow to see a list of variables

- Or click on `states1` to see the data in spreadsheet format

---
# Starting an editable document

- Two common documents 
  - Script: minimalist place to save code
  - R Markdown: allows you to fluidly combine prose and code to export to all sorts of things [video](https://rmarkdown.rstudio.com/lesson-1.html)

- We will use R Markdown documents exclusively in this course

- But let's start with editing a script seeing the difference between the two can be helpful
  
---

# Scripts

- I will now demonstrate how to open a script

--

> **Open a script**

--

- Purpose of scripts is to save code without caring about a finished document

- Scripts support two elements:
  - Code
  - Comments (requires hashtag `#` so R knows what is a comment)
  
```{r}
# This is a comment. Below is some arbitrary math.
((5+5)/2)^2
```

---
# Using a script

- R interprets each line of a script as a command to execute

- To execute a line, click enter

- I will now demonstrate executing code in a script

--

> **Type some arbitrary math in your script and execute**

---
# Objects

- Objects are anything you tell R to temporarily save during a session. A dataset, a number, a list of numbers, etc.

- You can save any object to your computer or delete it when you close R. Deleting is fine as long as you saved the code that created the object(s).

- You should already have an object `states1`

---
# Creating an object

> **Add `x <-` to the arbitrary math in your script like I did below. Execute.**

```{r}
x <- ((5+5)/2)^2
```

--

- The `<-` symbol is called the **assignment operator**. It tells R to save the result of the thing on the right side as the name on the left side.

- The keyboard shortcut for the assignment is `Alt-` on PC or `Option-` on Mac.

---
# Why save an object?

- We create an object whenever we want to use something for later or save it as a file.

- Suppose `x` is *super* interesting but your boss wants to know what `x` is squared.

- You don't need to recreate `x`. You can just square `x` like below.

```{r}
x^2
```

- This becomes very useful when working with multiple datasets

---
# Cleaning our environment

- By default, R keeps all the objects in our environment

- Over time, this becomes too much

--

- I will now demonstrate two options to clean our environment
  - Manually
  - Automatically via settings

--

> **I recommend you change your settings to this.**

---
# R Markdown

> **Close your script. Your choice whether to save it or not.**

--

- I will now demonstrate how to start a R Markdown (.Rmd) document.

--

> **Start a .Rmd document.**

- Let's review the template. It is informative.

---
# Knit

- Knit refers to taking the combination of prose and code and knitting it into a HTML, PDF, Word doc, etc.

- Should see a knit buttom at the top of the editor pane. Clicking it will knit your document.

  - Keyboard shortcut is `Cmd+Shift+K`

> **Knit your Rmd document. Name it "test"**

--

- Should see some new files in your file pane

- Let's compare the export to the Rmd document to learn some of how markdown works

---
# Cheatsheets

- The R Markdown cheatsheet is also informative

> **Go to Help->Cheatsheets->R Markdown Cheat Sheet**

---
# YAML header

- The YAML header tells R how to format the output

- We will learn additions later

> **Change the YAML to the following:**

```{r, eval=FALSE}
---
title:  'Test'
author: 'Your Name'
output: html_document
---
```

> **Knit to HTML again**

---
class: inverse, middle, center

# Let's create our first RLab and upload to eLC

---
# Starting RLab1

- I will now demonstrate adding another file to our project folder

--

> **Download the file "RLab1_Profile" from eLC and add to your "RLab1" project folder**

> **Rename the file "FirstName_LastName_Profile"**

> **Complete the instructions in the "RLab1_Profile" until you reach the section titled "Some of my R skills"**

---
# Adding code chunks

- Unlike a script, we must tell RStudio when to interpret code using a **code chunk**

- To add a code chunk, use the keyboard shortcut `Cmd+Opt+I`

- I will now demonstrate

--

> **Add a code chunk** 

> **Inside the code chunk, write some arbitrary math**

> **Assign the result to an object named `math`**

---
# Running a code chunk

- Several options to run code

- I will now demonstrate

- Simplest option is the keyboard shortcut `Cmd+Enter`

---
# Packages

- Many functions we want to use require packages.

- Your Rmd won't knit if you use code that requires a package that you haven't loaded **inside** of your Rmd.

--

> **Add another code chunk to your Rmd.**

> **Add the below code to it and run**

```{r, eval=FALSE}
glimpse(states1)
```

--

```{r, eval=FALSE}
Error in glimpse(states1) : could not find function "glimpse"
```

- This is because we haven't loaded the package to which `glimpse` belongs.

---
# Installing packages

- To load a package, it has to be installed on your computer. You only need to install a package once.

- If you haven't already, let's install the `tidyverse` package to our copy of R.

> **Run the following code in your console**

```{r, eval=FALSE}
install.packages("tidyverse")
```

--

- **Never** install packages inside of your Rmd. This causes R to install every time you knit. Only install in your console.

---
# Loading packages

- Once installed, now we can load a package whenever we need it. 

> **Add another code chunk to the bottom of your Rmd.**

> **Add this code to it and run.**

```{r, eval=FALSE}
library(tidyverse)
```

---
# Knitting in order

> **Now try running the `glimpse(states1)` code from before.**

- The result of the code should now be visible inside of your Rmd.

--

> **Now knit your Rmd.**

--

Error?! Let's investigate.

---
# Knitting in order

- R knits in the order of your code chunks. If you use a function that requires a package but you load the package afterwards, then R won't recognize the function when you knit.

--

> **Copy-and-paste the `library(tidyverse)` chunk above the `glimpse(states1)` chunk**

> **Knit to HTML again.**

---
# Upload to eLC

- Upload your Rmd file to the RLabs assignment dropbox on eLC.
