---
title: "PADP 7120 Data Applications in PA"
subtitle: "RLab 8: Reporting Regression Results"
author: "Alex Combs"
institute: "UGA | SPIA | PADP"
date: "October 21, 2020"
output:
  xaringan::moon_reader:
    css: 'mypres.css'
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      countIncrementalSlides: false
      highlightlines: true
---
# Objectives

By the end of this lab, you will have learned how to...

- Report regression results via tables and graphs in a realistic and accurate manner

```{r, include=FALSE}
library(tidyverse)
library(moderndive)
library(knitr)
library(carData)
library(stargazer) # for regression tables
library(jtools) # for regression visualizations
library(interactions) # for regression viz of interactions
```

---
# Set-up

- Start a new project and notebook
- Open your script from the R Chapter on regression with interactions (or the one I provided as answers)

- Copy and paste the script code into the notebook (remember to place the code inside code chunks)

---
# Set-up

Change YAML

```{r, eval=FALSE}
---
title: "RLab 8: Reporting Regression Results"
author: "Your Name"
output: 
  html_document:
    theme: spacelab
    df_print: paged
---
```

---
# Packages

```{r, eval=FALSE}
library(tidyverse)
library(moderndive)
library(knitr)
library(carData)
library(stargazer) # for regression tables
library(jtools) # for regression visualizations
```

- Remember to set `include=FALSE`

---
# Reporting Results

- You ran a regression that explains salaries.

- You are proud of your work and need to provide the results in a report or a presentation.

- Make sure the following code is in a code chunk. Knit to HTML.

```{r, eval=FALSE}
sal_mod <- lm(salary ~ sex + yrs.service + sex*yrs.service, data = Salaries)
get_regression_table(sal_mod)
```


---
# Reporting Results

- The tables you've produced so far are not typically used

```{r}
sal_mod <- lm(salary ~ sex + yrs.service + sex*yrs.service, data = Salaries)
```
```{r, eval=FALSE}
get_regression_table(sal_mod)
```
```{r, include=FALSE}
get_regression_table(sal_mod) %>% 
  kable()
```

---
# Reporting results

- Conventional regression results tables include:
  - estimate
  - standard error or t-statistic; never both
  - stars to indicate statistical significance based on p-value
  - Nicely formatted variable names
  
- Sometimes include confidence intervals

---
# Option 1: kable

- Add this code and knit to HTML again

```{r, eval=FALSE}
regtable.colnames <- c('Variable', 'Estimate', 
                       'SE', 'P-Value', 
                       'Lower 95%', 'Upper 95%')

get_regression_table(sal_mod) %>% 
  select(-statistic) %>% 
  kable(col.names = regtable.colnames, digits = 0)
```

---
# Option 2: stargazer

- Add this code. Include the chunk option `results='asis'`. Knit to HTML.

```{r, eval=FALSE}
stargazer(sal_mod,
          type="html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years of Service", "Male X Service"),
          keep.stat=c("n","rsq"),
          digits = 0,
          digits.extra = 3)
```

---
# Option 2: stargazer

- Change first line like so to include confidence intervals instead of standard errors. Knit to HTML.

```{r, eval=FALSE}
stargazer(sal_mod, ci=TRUE, ci.separator = "-",
          type="html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years of Service", "Male X Service"),
          keep.stat=c("n","rsq"),
          digits = 0,
          digits.extra = 3)
```

---
# Option 2: stargazer export to Word

- Add last line like so generate a new document named `sal_mod.htm` that you can open in Word. Knit to HTML.

```{r, eval=FALSE}
stargazer(sal_mod, ci=TRUE, ci.separator = "-",
          type="html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years of Service", "Male X Service"),
          keep.stat=c("n","rsq"),
          digits = 0,
          digits.extra = 3,
          out = "sal_mod.htm")
```

---
# Option 2: stargazer

- Also plays nicely with PDF provided you installed `tinytex` back in RLab2

```{r, eval=FALSE}
# First
install.packages('tinytex') # hit Enter
# Next
tinytex::install_tinytex() # hit Enter, this will take a few minutes
# if you ever need to uninstall TinyTeX, run tinytex::uninstall_tinytex()
```

---
# Option 2: stargazer

- Delete `type="html"` and table will be formatted for PDF export instead

```{r, eval=FALSE}
stargazer(sal_mod, ci=TRUE, ci.separator = "-",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years of Service", "Male X Service"),
          keep.stat=c("n","rsq"),
          digits = 0,
          digits.extra = 3)
```

---
class: inverse, center, middle

# Tables aren't great for presentations. We want to use graphs instead.

---
# Visualizing Regression Line

- We already know one option

.pull-left[
```{r, eval=FALSE}
ggplot(Salaries, 
       aes(x = yrs.service, 
           y = salary, 
           color = sex)) +
  geom_point(alpha = 0.6, size = 5) +
  geom_smooth(method = 'lm', se = FALSE, size = 2) +
  theme_minimal() +
  theme(text = element_text(size = 20))
```
]

.pull-right[
```{r, echo=FALSE, message=FALSE, out.width='99%'}
ggplot(Salaries, aes(x = yrs.service, y = salary, color = sex)) +
  geom_point(alpha = 0.6, size = 5) +
  geom_smooth(method = 'lm', se = FALSE, size = 2) +
  theme_minimal() +
  theme(text = element_text(size = 20))
```
]

---
# Visualizing Regression Line

- This is about as far as `ggplot` can go
- If we had one more variable in our regression, we wouldn't be able to visualize it this way because there is not an easy way to add it to the graph

---
# Visualizing Regression Line

- Add code

```{r, fig.align='center', fig.height=4}
effect_plot(sal_mod, pred = yrs.service, plot.points = TRUE)
```

---
# Visualizing Regression Line

```{r, fig.align='center', fig.height=4}
interact_plot(sal_mod, pred = yrs.service, modx = sex, plot.points = TRUE)
```

---
# Visualizing regression estimates

- Very common to visualize the estimates instead of the regression line. 

- Add following code and knit to HTML.

```{r, eval=FALSE}
plot_summs(sal_mod, 
           coefs = c("Male" = "sexMale", 
                     "Years of Service" = "yrs.service", 
                     "Male X Service" = "sexMale:yrs.service"))
ggsave('sal_mod_est.png')
```

```{r, echo=FALSE, message=FALSE, fig.align='center', fig.height=4}
plot_summs(sal_mod, 
           coefs = c("Male" = "sexMale", 
                     "Years of Service" = "yrs.service", 
                     "Male X Service" = "sexMale:yrs.service"))
```

