---
title: "PADP 7120 Data Applications in PA"
subtitle: "RLab 8: Reporting Regression Results"
author: "Alex Combs"
institute: "UGA | SPIA | PADP"
date: "October 21, 2020"
output:
  xaringan::moon_reader:
    css: 'mypres.css'
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      countIncrementalSlides: false
      highlightlines: true
---
# Objectives

By the end of this lab, you will have learned how to...

- Report regression results via tables and graphs in a realistic and accurate manner

```{r, include=FALSE}
library(tidyverse)
library(moderndive)
library(knitr)
library(carData)
library(stargazer) # for regression tables
library(jtools) # for regression visualizations
library(interactions) # for regression viz of interactions
```

---
# Set-up

- Start a new project and notebook

- Change YAML

```{r, eval=FALSE}
---
title: "RLab 8: Reporting Regression Results"
author: "Your Name"
output: 
  html_document:
    theme: spacelab
    df_print: paged
---
```

---
# Packages

- Remember to set `include=FALSE`

```{r, eval=FALSE}
library(tidyverse)
library(moderndive)
library(knitr)
library(carData)
library(stargazer) # for regression tables
library(jtools) # for regression visualizations
library(interactions) # for regression viz of interactions
```

---
# Revisiting Professor Salaries

- In R Chapter 22 (interactions), you ran the following model.

```{r}
sal_mod <- lm(salary ~ sex + yrs.service + sex*yrs.service, 
              data = Salaries)
```
```{r, eval=FALSE}
get_regression_table(sal_mod)
```
```{r, include=FALSE}
get_regression_table(sal_mod) %>% 
  kable()
```

- Add this code to your notebook

---
# Revisiting Professor Salaries

```{r, echo=FALSE, message=FALSE, fig.align='center', fig.height=6}
ggplot(Salaries, aes(x = yrs.service, y = salary, color = sex)) +
  geom_point(alpha = 0.6, size = 5) +
  geom_smooth(method = 'lm', se = FALSE, size = 2) +
  theme_minimal() +
  theme(text = element_text(size = 20))
```

- The data suggests we may want to model `yrs.service` using a quadratic

---
# Quadratic Salary Model

- Let's run a new model for `salary` that allows our regression to change direction once.

$$Salary = \beta_0 + \beta_1Male + \beta_2YrsServe + \beta_3YrsServe^2 + \epsilon $$

-Add the following code.

```{r}
sal_mod2 <- lm(salary ~ sex + yrs.service + I(yrs.service^2),
               data = Salaries)
```

---
# Reporting Results

- You are proud of your work and need to provide the results in a report or a presentation.

- Add following code. Knit to HTML.

```{r, eval=FALSE}
get_regression_table(sal_mod2)
```

---
# Reporting Results

- This table isn't great for an external audience

```{r, echo=FALSE}
get_regression_table(sal_mod2) %>% 
  kable()
```

---
# Reporting results

- Regression results tables usually include:
  - estimate
  - standard error or t-statistic; rarely both
  - stars to indicate statistical significance based on p-value
  - clear variable names
  - number of observations

- Sometimes goodness-of-fit measure like $R^2$  
- Sometimes include confidence intervals

---
# Option 1: kable

- Add heading "Kable"
- Add this code and knit to HTML again

```{r, eval=FALSE}
regtable.colnames <- c('Variable', 'Estimate', 
                       'SE', 'P-Value', 
                       'Lower 95% CI', 'Upper 95% CI')

get_regression_table(sal_mod2) %>% 
  select(-statistic) %>% 
  kable(col.names = regtable.colnames)
```

---
# Option 2: stargazer

- Add heading "Stargazer"
- Add this code. Include the chunk option `results='asis'`. Knit to HTML.

```{r, eval=FALSE}
stargazer(sal_mod2,
          type="html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years Served", "Years Served Sq."),
          keep.stat=c("n"), # reports only the number of observations
          digits = 0, # no decimals
          digits.extra = 3) #allows up to 3 decimals if rounding causes a 0
```

---
# Option 2: stargazer

- Change first line like so to include confidence intervals instead of standard errors.
- Knit to HTML.

```{r, eval=FALSE}
stargazer(sal_mod2, ci=TRUE, ci.separator = " : ",
          type="html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years Served", "Years Served Sq."),
          keep.stat=c("n"),
          digits = 0,
          digits.extra = 3)
```

---
# Option 2: stargazer

- Add last line like so to generate a new document named `sal_mod2.htm` that you can open in Word. Knit to HTML.

```{r, eval=FALSE}
stargazer(sal_mod2, ci=TRUE, ci.separator = " : ",
          type="html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years Served", "Years Served Sq."),
          keep.stat=c("n"),
          digits = 0,
          digits.extra = 3,
          out = "sal_mod2.htm")
```

---
# Option 2: stargazer

- Also plays nicely with PDF provided you installed `tinytex` back in RLab2

```{r, eval=FALSE}
# First
install.packages('tinytex') # hit Enter
# Next
tinytex::install_tinytex() # hit Enter, this will take a few minutes
# if you ever need to uninstall TinyTeX, run tinytex::uninstall_tinytex()
```

---
# Option 2: stargazer

- Delete `type="html"` and table will be formatted for PDF export instead

```{r, eval=FALSE}
stargazer(sal_mod2, ci=TRUE, ci.separator = " : ",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years Served", "Years Served Sq."),
          keep.stat=c("n"),
          digits = 0,
          digits.extra = 3)
```

---
class: inverse, center, middle

# Tables aren't great for presentations. We want to use graphs instead.

---
# Visualizing Regression Line

- We already know one option using `ggplot`

.pull-left[
```{r, eval=FALSE}
ggplot(Salaries, 
       aes(x = yrs.service, 
           y = salary, 
           color = sex)) +
  geom_point(alpha = 0.6, size = 5) +
  geom_smooth(method = 'lm', se = FALSE, size = 2) +
  theme_minimal() +
  theme(text = element_text(size = 20))
```
]

.pull-right[
```{r, echo=FALSE, message=FALSE, out.width='99%'}
ggplot(Salaries, aes(x = yrs.service, y = salary, color = sex)) +
  geom_point(alpha = 0.6, size = 5) +
  geom_smooth(method = 'lm', se = FALSE, size = 2) +
  theme_minimal() +
  theme(text = element_text(size = 20))
```
]

---
# Visualizing Regression Line

- At most, `ggplot` can easily visualize regression models that include:
  - **ONE** numerical explanatory variable
  - **ONE** categorical explanatory variable (interacted with numerical or not)
  
- The quadratic is difficult to add
- If we decided to control for one more variable, our graph would not be showing the relationship between salary and years of service **controlling for that other variable**

---
# Visualizing Regression Line

- `effect_plot` from the `jtools` package makes this very easy. Add code:

```{r, fig.align='center', fig.height=5}
effect_plot(sal_mod2, pred = yrs.service, plot.points = TRUE)
```

---
# Visualizing Regression Line

- If we want to include Male, `interact_plot` from `interactions` package is useful. 
- Works the same had we interacted `sex` with `yrs.service`.

```{r, fig.align='center', fig.height=5, warning=FALSE}
interact_plot(sal_mod2, pred = yrs.service, modx = sex, plot.points = TRUE)
```

---
# Visualizing regression estimates

- Very common to visualize the estimates instead of the regression line. 

- Add following code and knit to HTML.

```{r, eval=FALSE}
plot_summs(sal_mod2, 
           coefs = c("Male" = "sexMale", 
                     "Years Served" = "yrs.service", 
                     "Years Served Sq." = "I(yrs.service^2)"))
ggsave('sal_mod2_est.png') # saves most recent graph to separate file
```

```{r, echo=FALSE, message=FALSE, fig.align='center', fig.height=3.5}
plot_summs(sal_mod2, 
           coefs = c("Male" = "sexMale", 
                     "Years Served" = "yrs.service", 
                     "Years Served Sq." = "I(yrs.service^2)"))
```

---
# Multiple models

- We have two models `sal_mod` and `sal_mod2` we may want to report together
- Add code and knit to HTML

```{r, eval=FALSE}
stargazer(sal_mod, sal_mod2, ci=TRUE, ci.separator = " : ",
          type = "html",
          title="Regression Results",
          dep.var.labels=c("Salary"),
          covariate.labels=c("Male", "Years Served","Males X Years Served", "Years Served Sq."),
          keep.stat=c("n"),
          digits = 0,
          digits.extra = 2)
```

---
# Multiple Models

- Add code. Knit to HTML.

```{r, eval=FALSE}
plot_summs(sal_mod, sal_mod2, 
           coefs = c("Male" = "sexMale", 
                     "Years Served" = "yrs.service", 
                     "Male X Years Served" = "sexMale:yrs.service",
                     "Years Served^2" = "I(yrs.service^2)"))
```

---
# Save and upload to eLC