---
title: "PADP 7120 Data Applications in PA"
subtitle: "RLab 5: Visualizations"
author: "Alex Combs"
institute: "UGA | SPIA | PADP"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  xaringan::moon_reader:
    css: 'mypres.css'
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      countIncrementalSlides: false
      highlightLines: true
---
# Objectives

- Create graphs that are appropriate based on data and intended message

- Make some basic improvements to default graphs

```{r, include=FALSE}
library(knitr)
```


---
# Setup

> **Start a new project named "rlab5".**
  
> **Start a R Markdown document**

> **Keep the setup code chunk at the top but delete the rest of the template. Change echo to FALSE.**

> **Change YAML to:**

```{r, eval=FALSE}
---
title: "RLab5: Data Viz"
author: "Your Name"
output:
  html_document:
    df_print: paged
---
```

---
# Setup

> **In the top "setup" code chunk, load the following packages.**

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(lubridate)
```

---
# County COVID Data

- Let's examine the latest COVID data for every U.S. county

> **Add and run the following code**

```{r}
covid <- fread('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
```

- These data include *cumulative* counts (not daily) of COVID cases and deaths

---
# Some Wrangling

```{r, echo=FALSE}
glimpse(covid)
```

- Let's focus on Georgia

> **Create a new dataset called `covid_ga` that contains only counties in Georgia**

```{r, include=FALSE}
covid_ga <- covid %>% 
  filter(state == 'Georgia')
```

---
# Variable Type Conversions

```{r, echo=FALSE}
glimpse(covid_ga)
```

- We need to tell R to recognize `date` as dates instead of text

> **Add and run following code that tells R `date` is organized as `y`ear-`m`onth-`d`ay.**

```{r}
covid_ga$date <- ymd(covid_ga$date)
```

---
# Variable type conversions

```{r, echo=FALSE}
glimpse(covid_ga)
```

- Now date is stored as a date instead of character

---
# Quick Summary

> **Use `summary` to get a quick look at summary stats. Suppress output.**

```{r, include=FALSE}
summary(covid_ga)
```

--

- Note the latest date with data is yesterday. We will want to focus on the most recent counts, so let's save yesterday's date for easy reference.

> **Add and run following code.**

```{r}
yesterday <- '2021-09-10'
```

- Now we can add `yesterday` to our code any time we want to filter conditional on the most recent data, and we only need to update this code to update any graph we make.

---
# Viz 1

- Suppose we are asked for the distribution of cumulative county cases in GA for the most recent available date.

- What kind of graph should we use?

--

> **In a new code chunk, write code to generate the correct graph.**

---
# Viz 1

```{r, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = cases)) +
  geom_histogram()
```

---
# Viz 2

> **In a new code chunk, write code to generate another histogram for cumulative deaths during the most recent date.**

--

```{r, echo=FALSE, message=FALSE, fig.width=8, fig.height=6, fig.align='center'}
covid_ga %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = deaths)) +
  geom_histogram()
```

---
# Viz 3

- Suppose we are asked to visualize the distribution of cumulative county deaths during the most recent date that includes visual markers for descriptive statistics and outliers

- Which graph should we use?

--

> **In a new code chunk, write code to generate the correct graph.**

---
# Viz 3

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga %>% 
  filter(date == yesterday) %>%
  ggplot(aes(x = deaths)) +
  geom_boxplot()
```

---
# County comparison

- Suppose we want to focus on five counties: Fulton, Cobb, DeKalb, Gwinnett, and Clarke.

> **Copy and paste below code to create a new data set that includes only these counties.**

```{r}
covid_ga_5 <- covid_ga %>% 
  filter(county %in% c('Fulton', 'Cobb', 'DeKalb', 'Gwinnett', 'Clarke'))
```

```{r, eval=FALSE}
covid_ga <- covid_ga %>% 
  mutate(county5 = fct_other(county, 
                             keep = c('Fulton', 'Cobb', 'DeKalb', 'Gwinnett', 'Clarke')))
```

---
# Viz 4

- Suppose we are asked to compare cumulative cases as of the most recent day in the data across our 5 counties.

- What kind of visualization could/should we use?

--

- Bar graph. But both `geom_bar` or `geom_col` make bar graphs. Which should we use?

--

- Counts are already aggregated. Therefore, we should use `geom_col`.

> **Generate a bar graph for cumulative cases during the most recent date across our 5 counties. Make another for deaths.**

---
# Viz 4

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col()
```

---
# Viz 5

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = deaths)) +
  geom_col()
```

---
# Viz 6 & 7

- Suppose we want to compare the trends in cases and deaths for our 5 counties over the time span of our data.

- Which graph should we use?

--

> **Generate the appropriate graph each for cases and deaths over the time span in the data for our 5 counties. Give each county its own line. To do so, you will need to include a third `aes`thetic in addition to `x` and `y`.**

---
# Viz 6

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  ggplot(aes(x = date, y = cases, color = county)) +
  geom_line()
```

---
# Viz 7

```{r, echo=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  ggplot(aes(x = date, y = deaths, color = county)) +
  geom_line()
```

---
# What's the Problem?

- All of our comparisons between counties have been somewhat flawed or potentially misleading. Why?

--

- We haven't adjusted for county population. If our goal is to compare spread of COVID-19 across counties, we need to account for the underlying population.

--

- Therefore, we need to compute rates; cases and/or deaths *per* some standard number of people. But these data do not include county population.

- We need more data to add to these data. We'll learn how to combine data in the next lab.

---
class: inverse, center, middle

# Let's cover some ways to make basic improvements to our graphs.

---
# Basic improvements

- Titles and axis labels
- Themes
- Text size

---
# Improving Viz 4

> **Find your code for Viz 4 (bar graph of cases for 5 counties) and copy it to a new code chunk.**

---
# Titles and axis labels

- Titles and all axis labels are controlled by `labs`

> **Add the lines below to your graph.**

```{r, eval=FALSE, message=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases', #<<
       x = '',#<<
       y = 'Number of cases')#<<
```

---
# Titles and axes labels

```{r, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases', #<<
       x = '',#<<
       y = 'Number of cases')#<<
```

---
# Themes

- R comes with numerous default themes to change the look of graphs

```{r, eval=FALSE}
theme_classic()
theme_minimal()
theme_void()
theme_bw()
# and more
```

---
# Themes

> **Add the new line below to your current code and re-run to see what it did.**

```{r, eval=FALSE, message=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases',
       x = '',
       y = 'Number of cases') +
  theme_minimal() #<<
```

---
# Themes

```{r, echo=FALSE, message=FALSE, fig.width=10, fig.height=8, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases',
       x = '',
       y = 'Number of cases') +
  theme_minimal() #<<
```

---
# Text size

- After using `theme_whatever`, we can control any of the "ink" via `theme()`

> **Add new lines below to your current code and re-run to see what it did**

```{r, eval=FALSE}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases',
       x = '',
       y = 'Number of cases') +
  theme_minimal() +
  theme(axis.text=element_text(size=16), #<<
        title = element_text(size=16)) #<<
```

---
# Text size

```{r, echo=FALSE, message=FALSE, fig.width=9, fig.height=7, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases',
       x = '',
       y = 'Number of cases') +
  theme_minimal() +
  theme(axis.text=element_text(size=16),
        title = element_text(size=16))
```

---
# Skill Check

> **Use labels, themes, and text size to improve Viz 7**

```{r, include=FALSE, fig.width=9, fig.height=7, fig.align='center'}
covid_ga_5 %>% 
  filter(date == yesterday) %>% 
  ggplot(aes(x = county, y = cases)) +
  geom_col() +
  labs(title = 'Total COVID Cases Among Selected Georgia Counties',
       x = '',
       y = 'Case Count',
       color = 'County') +
  theme_minimal() +
  theme(axis.text=element_text(size=16),
        title = element_text(size=16))
```

---
# Upload lab

- Upload your Rmd to eLC.
