---
title: "PADP 7120 Data Applications in PA"
subtitle: "RLab 7: Polished Visualizations"
author: "Alex Combs"
institute: "UGA | SPIA | PADP"
date: "September 23, 2020"
output:
  xaringan::moon_reader:
    css: 'mypres.css'
    nature:
      beforeInit: "macros.js"
      highlightStyle: github
      countIncrementalSlides: false
      highlightlines: true
---
# Objectives

By the end of this lab, you will have learned how to...

- Polish graphs to share with an audience  

---
# Start a new project
- Using the file directory on your computer
  - Go to the parent folder that contains your RLabs folders
  - Create a new folder named RLab7

--

- In RStudio
  - Create new project
  - Select Existing Directory
  - Click Browse and select the RLab7 folder you just created
  - Click Create Project
  
---
# Start New Notebook

- Change YAML

```{r, eval=FALSE}
---
title: "RLab 7: Polished Visualizations"
author: "Your Name"
output: 
  html_document:
    theme: spacelab
    df_print: paged
---
```

---
# Packages

- Load the following packages, suppressing all code and output.

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(lubridate)
```

---
# Copy and Paste Previous Labs

- Open your notebooks from RLabs 5 & 6

- From RLab 5 or 6, copy and paste the code needed to generate the `covid_ga` dataset

```{r}
covid <- fread('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')

covid_ga <- covid %>% 
  filter(state == 'Georgia')

covid_ga$date <- ymd(covid_ga$date)

today <- '2020-08-16'
```

---
# Copy and Paste Previous Labs

```{r}
covid_ga <- covid_ga %>% 
  group_by(county) %>% 
  arrange(date) %>% 
  mutate(day_cases = cases - lag(cases),
         day_deaths = deaths - lag(deaths)) %>% 
  ungroup()
```

---
# Copy and Paste Previous Labs

- Copy `ga_county_pop.csv` from your RLab 6 folder and paste into your RLab 7 folder
- From RLab 6 notebook, copy and paste the code that imported and left joined `ga_county_pop`

```{r, eval=FALSE}
ga_county_pop <- read_csv('ga_county_pop.csv')
```

```{r, include=FALSE}
ga_county_pop <- read_csv('labs_files/ga_county_pop.csv')
```

```{r}
covid_ga <- covid_ga %>% 
  left_join(ga_county_pop, by = 'county')

covid_ga$county <- as.factor(covid_ga$county)
```

---
# Copy and Paste Previous Labs

- Finally, copy and paste all the code used to make the visualizations from RLab 5

---
# Brief Aside About Replication

- In a real scenario, we would probably just update the notebook we've been working on since RLab 5

- But suppose we needed to start a new project that involved similar code

- We don't have to start from scratch or reverse engineer point-and-clicks

- Once we have code that works, we have a template for future projects

---
# Rerun Code

- If you haven't done so already, run the code to get the `covid_ga` dataset that includes county population

---
# Adjusting for Population

- We decided in RLab5 that it is not right to compare counts of COVID cases and/or deaths across counties of different size populations

--

- We need to compute counts *per* capita or some standard number of people, such as 100,000

- Let's create four new variables, `cases_rate`, `deaths_rate`, `day_cases_rate`, and `day_deaths_rate`, using per 100,000 people.

--

Before trying in R, think through and describe the steps we need to take to obtain cases or deaths per 100,000 given our data in `covid_ga`.

---
# Adjusting for Population

  - We need to create a new variable for population in 100,000s by dividing population by 100,000

--

  - Then, create the rate variables by dividing cases and deaths by this new measure of population in 100,000s

--

  - Now that we know the plan, try executing it in R, overwriting the `covid_ga` dataset

---
# Adjusting for Population

```{r}
covid_ga <- covid_ga %>% 
  mutate(pop100thou = population/100000) %>% 
  mutate(cases_rate = cases/pop100thou, 
         deaths_rate = deaths/pop100thou,
         day_cases_rate = day_cases/pop100thou, 
         day_deaths_rate = day_deaths/pop100thou)
```

---
# Selected Counties

- First, quick summary

```{r, results='hide'}
summary(covid_ga)
```

- Now we are ready to recreate the dataset that includes only the five counties we selected from before, which will now have all the changes we made to `covid_ga`

```{r}
ga_counties <- c('Fulton', 'Cobb', 'DeKalb', 'Gwinnett', 'Clarke')

covid_ga_five <- covid_ga %>% 
  filter(county %in% ga_counties)
```

---
# Common Viz Adjustments

- Non-data ink
  - Labels (title and axes)
  - Themes
  - Axes & scales (tick marks, commas, dollars, log scale)

--

- Geometric object
  - Color
  - Shape/line type & size
  - Transparency
  - Reference lines

---
# Viz 1 & 2

- Suppose we are asked for the distribution of county cases for the most recent available date.
- What kind of graph should we use?

- A histogram
- Generate a histogram for cases during the most recent date
- Suppress code and any messages
- Generate another histogram for deaths during the most recent date

---
# Viz 1

- Let's improve this graph by adding a title and labels as well as change the theme to one of R's default themes

```{r, message=FALSE, fig.width=6, fig.height=4}
covid_ga %>% 
  filter(date == today) %>% 
  ggplot(aes(x = cases)) +
  geom_histogram()
```

---
# Viz 1

```{r, message=FALSE, fig.width=5, fig.height=3}
covid_ga %>% 
  filter(date == today) %>% 
  ggplot(aes(x = cases)) +
  geom_histogram() +
  labs(title = 'Total County Cases Today',
       x = 'Total Cases',
       y = 'Counties') +
  theme_minimal()
```

---
# Viz 1

- Let's fix the skew using log scale

```{r, message=FALSE, fig.width=5, fig.height=3}
covid_ga %>% 
  filter(date == today) %>% 
  ggplot(aes(x = cases)) +
  geom_histogram() +
  labs(title = 'Total County Cases Today',
       x = 'Total Cases',
       y = 'Counties') +
  scale_x_log10() +
  theme_minimal()
```

---
# Viz 1

- Let's change the color and outline of the bins of the histogram

```{r, message=FALSE, fig.width=5, fig.height=3}
covid_ga %>% 
  filter(date == today) %>% 
  ggplot(aes(x = cases)) +
  geom_histogram(fill = 'steelblue', color = 'white') +
  labs(title = 'Total County Cases Today',
       x = 'Total Cases',
       y = 'Counties') +
  scale_x_log10() +
  theme_minimal()
```

---
# Viz 1: Adding Comma Separator

```{r, message=FALSE}
library(scales)
```

```{r, message=FALSE, fig.width=5, fig.height=3}
covid_ga %>% 
  filter(date == today) %>% 
  ggplot(aes(x = cases)) +
  geom_histogram(fill = 'steelblue', color = 'white') +
  labs(title = 'Total County Cases Today',
       x = 'Total Cases',
       y = 'Counties') +
  scale_x_log10(labels = comma) +
  theme_minimal()
```

---
# Viz 2

- On your own, make the same adjustments for the histogram of total county deaths

---
# Viz 3

- Suppose we are asked for the distribution of county deaths during the most recent date that includes visual markers along key points of the distribution
- Which graph should we use?

- A box-and-whiskers plot
- Generate a box plot for county deaths during the most recent date. Suppress the code.

---
# Viz 3

.pull-left[
```{r, fig.width=5, fig.height=4}
covid_ga %>% 
  filter(date == today) %>%
  ggplot(aes(x = deaths)) +
  geom_boxplot()
```
]

.pull-right[
- Common issue with box plots is that one axis doesn't mean anything.
- In this case, we should remove the tick marks, line, and text for the y axis
]

---
# Viz 3: Remove axis elements

```{r, fig.width=5, fig.height=4}
covid_ga %>% 
  filter(date == today) %>%
  ggplot(aes(x = deaths)) +
  geom_boxplot() +
  theme(axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.text.y = element_blank())
```

---
# Viz 3: Change background and color

```{r, fig.width=5, fig.height=4}
covid_ga %>% 
  filter(date == today) %>%
  ggplot(aes(x = deaths)) +
  geom_boxplot(fill = 'steelblue') +
  theme_classic() +
  theme(axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.text.y = element_blank())
```

---
# Viz 3: Title and axis label

```{r, fig.width=5, fig.height=3}
covid_ga %>% 
  filter(date == today) %>%
  ggplot(aes(x = deaths)) +
  geom_boxplot(fill = 'steelblue') +
  labs(title = 'Total County COVID Deaths as of Today',
       x = 'Total Deaths') +
  theme_classic() +
  theme(axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.text.y = element_blank())
```

---
# Viz 4

- Rather than total cases or deaths for all counties as of the most recent date, let's focus on the distribution daily counts for a single county.

- Generate a box plot for Clarke county daily cases during all available dates

- Make the same adjustments to this box plot as we did the last one

---
# Viz 4

```{r, fig.width=5, fig.height=3}
covid_ga %>% 
  filter(county == 'Clarke') %>%
  ggplot(aes(x = cases)) +
  geom_boxplot(fill = 'steelblue') +
  labs(title = 'Daily COVID Cases for Clarke County',
       subtitle = 'January 21 - September 30',
       x = '') +
  theme_classic() +
  theme(axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        axis.text.y = element_blank())
```

---
# Viz 5

- Suppose we are asked to compare cases for the most recent day in the data across our 5 counties
- Cases and deaths are already aggregated. Therefore, we should use `geom_col`.

- Generate a bar graph for cases *per 100,000* during the most recent date across our 5 counties

---
# Viz 5

```{r, fig.width=7, fig.height=5}
covid_ga_five %>% 
  filter(date == today) %>% 
  ggplot(aes(x = county, y = cases_rate)) +
  geom_col()
```

---
# Viz 5: Reorder bars

- We typically want heights of bars to be in ascending or descending order

- Let's reorder to descending

```{r, fig.width=6, fig.height=4}
covid_ga_five %>% 
  filter(date == today) %>% 
  ggplot(aes(x = reorder(county, -cases_rate), y = cases_rate)) +
  geom_col()
```

---
# Viz 5: Other Adjustments

```{r, fig.width=5, fig.height=3}
covid_ga_five %>% 
  filter(date == today) %>% 
  ggplot(aes(x = reorder(county, -cases_rate), y = cases_rate)) +
  geom_col(fill = 'red', color = 'white') +
  scale_y_continuous(labels = comma) +
  labs(title = 'COVID Cases Per 100,000 Today',
       x = 'County',
       y = 'Cases per 100,000') +
  theme_minimal()
```

---
# Viz 6

- Suppose we want to show the trends in cases and deaths for our 5 counties

- Generate a line graph over the time span in the data for our 5 counties.
- Give each county its own line.
- Suppress the code.

- First, let's change this graph to use cases per 100,000 people

---
# Viz 6

```{r, fig.width=7, fig.height=5}
covid_ga_five %>% 
  ggplot(aes(x = date, y = cases_rate, color = county)) +
  geom_line()
```

---
# Viz 6: Change line type and size

```{r, fig.width=7, fig.height=5}
covid_ga_five %>% 
  ggplot(aes(x = date, y = cases_rate, color = county)) +
  geom_line(linetype = 'dashed', size = 4)
```

---
# Viz 6: Readjust

- Those changes aren't improvements, let's stick with the defaults

- Let's change the title of the legend

```{r, fig.width=6, fig.height=4}
covid_ga_five %>% 
  ggplot(aes(x = date, y = cases_rate, color = county)) +
  geom_line() +
  labs(legend = 'County')
```